Linear variation and an optimization of algorithms for connected components labeling in a binary image

### 2016-02-06 paper/paper.tex
Finish SUF subsubsubsubsubsection

@@

Like most approaches, the algorithm contains has and outer loop over all pixels
of the image.
Each time an object pixel that is not labeled yet is encountered, DFS traversal
is launched from that pixel.
The traversal labels all the object pixels of the same CC, launching recursively
from all unlabeled neighbors of current pixel.

@@

Another general approach for graph CC labeling which is more popular in image
labeling is using some kind of a data structure to maintain disjoint sets of pixels
that are considered to be in the same CC.
The structure is commonly called \texttt{Union-Find}, as the two essential operations
it must support are \texttt{Union}, which is used to union some of the disjoint sets,
and \texttt{Find}, which is used to find the representative element of the set
some element belongs to, so that we can always tell if two elements are in the same
set.

@@

>>>>

As each of these methods, omitting DTSUF could be used with both condensed and
non-condensed images, for convenience we denote height and width of the image
on input of each algorithm as $H$ and $W$.

Also to avoid repeating ``pixel or cell'' we will just use the word ``pixel''
meaning the appropriate raster for non-condensed and condensed images.
When we mention neighbors of some pixel, we mean pixels that are directly connected
to it in terms of appropriate connectivity.

@@

Like most approaches, the algorithm has an outer loop over all pixels of the image.
Each time an object pixel (non-empty cell, in case it is actually a cell)
which is not labeled yet is encountered, DFS traversal is launched from that pixel.
The traversal labels all the object pixels of the same CC, launching recursively
from all unlabeled neighbors of current pixel.

@@

Another general approach for graph CC labeling which is more popular in image
labeling is using some kind of a data structure to maintain disjoint sets of pixels
that are considered to be in the same CC.

The structure is commonly called Union-Find, as the two essential operations
it is supposed to support are Union, which is used to merge some two of the
disjoint sets in one, and Find, which is used to find the representative element
of the set some element belongs to, so that we can always tell if two elements are
in the same set.
The structure can be implemented so that the amortized time complexity for each
operation is $O(\alpha(HW))$, where $\alpha$ is the inverse Ackermann function,
and $\alpha(HW) < 5$ for all remotely practical image sizes. % TODO links

The approach that is named SUF (Scan plus Union-Find) in this paper follows.

Pixels are processed in the natural order: from top to bottom and from
left to right, so each row is fully processed before all rows after it.
So for each object pixel we encounter there are at most 4 nearby pixels
that were already processed, see \autoref{fig:suf:neighbors}.

At first each object pixel is considered to belong to its own CC, so right before
considering the neighbors we increment the CC counter.
Moreover, it is associated with its own disjoint set in the Union-Find structure.
More specifically, object pixel at position $x, y$ is associated with
the set number $xW + y$ to guarantee absence of collisions.

The possible neighbors are processed in the order shown on \autoref{fig:suf:neighbors},
and for each neighbor that is object pixel, the Union operation is performed
on the two sets containing the current and the neighbor pixel.
In case the Union operation was successful, which means that the pixels
had belonged previously to different sets and we have just found a way to merge two
CC, the CC counter is decremented.

\begin{figure}
  \centering
  \begin{tikzpicture}
    \foreach \x/\y [count=\i from 1] in {0/1,1/1,2/1,0/0} {
      \node[whitebox] at (\x,\y) {\i};
    }
    \node[graybox] at (1, 0) {};
    \node[whitebox] at (1, 0) {};
  \end{tikzpicture}
  \caption{Neighbor mask for SUF algorithm family.}
  \label{fig:suf:neighbors}
\end{figure}

Note that this algorithm outputs not the matrix of final labels, but only the number
of CC.
This is the place in which it differs from some common two-pass approaches for CC % TODO references
labeling, as it is an algorithm not for labeling, but just for counting CC.
The latter is exactly what is needed for the aforementioned linear variation problem.

Also we have to mention that for performance reasons the Union-Find internal structures
have to be allocated ahead to have the size of at least $HW$ to guarantee that
each object can be easily associated with unique starting set given only its
coordinates. So memory consumption of SUF is linear in the number of pixels.

@@

using desision trees suggested by Wu, Otoo and Suzuki

\subsection{Condensation types}

bla-bla about 2x2 in memory and as a view
